Stack Safety is a Security Property

### 2021-04-15 tex/main.tex
Some polishing and comments.

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
%\apt{maybe: ``Stack safety is a testable security property''?}\bcp{That
%  would sacrifice a lot of oomph for a little bit of precision, IMO.  Prefer
%  to leave it out of the title but hit it strongly in the abstract.}%
What is stack safety? The phrase is associated with a variety of compiler,
run-time, and hardware mechanisms for guarding against corruption of stack
memory.  But these mechanisms typically lack rigorous specifications,
relying instead on informal descriptions and specific examples of the bad behaviors that they prevent.

We propose formal criteria for stack safety, using the concepts of
language-based security: a combination of an integrity property (``each
caller's stack frame is held invariant by the callee'') and a confidentiality
property (``the callee's behavior is insensitive to differences in the
caller's state'').
% These properties are sensible even in the presence of
% abnormal control flow.\apt{Is this important enough to mention here?}
% BCP: I think not in the abstract, but maybe in the intro.
% We further apply\apt{what does this mean? we didn't
%  invent it} a straightforward notion of
% ``well-bracketed control flow'' that restricts control flow to the
% typical pattern of matched calls and returns.
%
We argue that these criteria match common intuitions by (1) showing that
they are violated by a variety of intuitively unsafe examples and (2)
using property-based random testing to check that appropriate variants of
the criteria are actually enforced by the stack-safety ``micro-policies''
proposed by~\citet{DBLP:conf/sp/RoesslerD18} (and {\em not} enforced by
broken variants).  Specifically, Roessler
and Dehon's ``eager'' micro-policy, which catches violations as early as
possible, enforces a basic ``stepwise'' versions of our confidentiality
and integrity properties. Their more performant ``lazy'' micro-policy
does not, but corresponds to weaker, more extensional ``observational'' variants
of the
properties.  While we do not rule out other enforcement mechanisms, our
definitions are strongly informed by micropolicy systems, especially
our attacker model, which is stronger than other enforcement mechanisms
can enable.

@@

The call stack is a perennial target for low-level attacks
based on memory-safety vulnerabilities, leading to consequences
that range from leakage or corruption of private stack data to
control-flow hijacking. To deter or detect such attacks, a profusion of
software and hardware protections have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b,Tsampas+19,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\bcp{Yes, going back to MIT days---we should
  include several more of these, if only to give readers the impression that
this is a well-studied mechanism (so formalizing its protections is
useful).}
\fi
%
The protections offered by such mechanisms are commonly described in terms
of concrete examples of attacks that they can prevent, such as corruption
of return addresses, buffer overflows, or use of uninitialized variables,
leaving a more abstract characterization to the reader's intuition.
But these mechanisms can be quite intricate
and often require trading off protection and efficiency.
Thus, it would be useful to have a precise, generic, and formal
specification for stack
safety, both as a basis for comparing the security claims of different
enforcement techniques and for validating more rigorously that these claims
are met by particular implementations.

%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
The most abstract characterization stack safety of which we are aware is in
recent work by \citet{Skorstengaard+19}, which defines stack safety as the conjunction of two
properties:
{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
Informally, LSE says that the contents of
a caller's stack frame are not read and written while its callees (and their callees, etc.) are
executing, and WBCF says that callees always
return to the point just after their call (if they return at all).
\citet{Skorstengaard+19} formalize these properties by defining an idealized
machine in which they hold {\em by construction};
they then show that an enforcement
mechanism (in their case, hardware capabilities) guarantees stack safety by
proving that an ordinary machine with enforcement can simulate the idealized machine.
%an enforcement mechanism (in their case, hardware
%capabilities)
%can then be shown to guarantee stack safety by exhibiting a fully abstract
%embedding function
%from the idealized machine to an ordinary machine equipped with this
%enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
%  understand that sentence.}
%%
%% While this work represents an important step in the right direction, we believe
%% there is still room to craft an operational characterization
%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
%% In other words, we seek formal criteria that can be applied
%% to a mechanism for protecting stack-based low-level programs to judge
%% whether or not it actually ``guarantees stack safety.''
%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
%% confused by the references to these types of properties, yet to be introduced?}
%% We begin with a strong
%% local state encapsulation property that closely models the behavior of eager
%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
%% is too strong to characterize stack safety; reasonable enforcement mechanisms
%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
%% to support this technique and argue that the new property, defined relative to an
%% arbitrary notion of observation. This property is still strong enough to prevent
%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
%% other reasonable enforcement mechanisms can enforce it.

In this work, we also distinguish between protecting data (like LSE) and
control flow (with WBCF), and formalize a data protection property
based on the technical framework of language-based security~\citep{??}.
Our core properties are \emph{stepwise stack integrity} and
\emph{stepwise stack confidentiality}. WBCF is a desireable property, but
we can express safety in an information-flow sense without reference to it,
even safety from attacks that hijack control flow.

@@

\section{Key Ideas, By Example}
\label{sec:running-example}
\colorlet{lgray}{gray!40}
\colorlet{lred}{red!40}
\colorlet{lblue}{blue!20}

@@

\definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep}
if their adjacent pairs change together. We will represent this relation with
\(\machT \doteq \nachT\). \bcp{People may not be familiar with the notation}
\sna{It's a new notation, we use this enough that it seems worth picking an
  evocative symbol.} 
We define it as a coinductive relation:

@@

\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \outside\).
\(\MPCT\) corresponds to the trace of a subroutine from its call to its return.
Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
{\em stepwise stack confidentiality} if for all such \(\MPCT\) and \(\NPCT\),
\(\MPCT \doteq \NPCT\).

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
%\apt{maybe: ``Stack safety is a testable security property''?}\bcp{That
%  would sacrifice a lot of oomph for a little bit of precision, IMO.  Prefer
%  to leave it out of the title but hit it strongly in the abstract.}%
What does ``stack safety'' mean, exactly? The phrase is associated with a
variety of compiler,
run-time, and hardware mechanisms for guarding against corruption of stack
memory.  But these mechanisms typically lack precise specifications,
relying instead on informal descriptions and specific examples of bad
behaviors that they prevent.

We propose formal criteria for stack safety, using the concepts of
language-based security: a combination of an integrity property (``the private
state in each caller's stack frame is held invariant by the callee'')
and a confidentiality property (``the callee's behavior is insensitive to the
caller's private state'').
% These properties are sensible even in the presence of
% abnormal control flow.\apt{Is this important enough to mention here?}
% BCP: I think not in the abstract, but maybe in the intro.
% We further apply\apt{what does this mean? we didn't
%  invent it} a straightforward notion of
% ``well-bracketed control flow'' that restricts control flow to the
% typical pattern of matched calls and returns.
%
We argue that these criteria match common intuitions by (1) showing that
they are violated by a variety of intuitively unsafe examples and (2)
using property-based random testing to check that appropriate variants of
the criteria are actually enforced by the stack-safety ``micro-policies''
proposed by~\citet{DBLP:conf/sp/RoesslerD18} (and {\em not} enforced by
broken variants).  Specifically, Roessler
and Dehon's ``eager'' micro-policy, which catches violations as early as
possible, enforces a basic ``stepwise'' versions of our confidentiality
and integrity properties. Their more performant ``lazy'' micro-policy
does not, but it does (almost) correspond to weaker, more extensional
variants of the properties.
%
\BCP{I still don't think the next sentence really nails it.  And (sigh) I am
  still not really clear yet on whether our definition applies / might apply
  to other enforcement mechanisms or not.  Saying that our attacker model is
  stronger does not help---that should make it easier for us to describe
  their protections, not harder.  So I don't understand what we're saying
  here.}%
%
While we do not rule out other enforcement mechanisms, our
definitions are strongly informed by micropolicy systems, especially
our attacker model, which is stronger than other enforcement mechanisms
can enable.

\bcp{We should decide between ``criteria'' and ``properties''...}

@@

The call stack is a perennial target for low-level attacks, leading to
consequences
from leakage or corruption of private stack data to
control-flow hijacking. To deter or detect such attacks, a profusion of
software and hardware protections have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b,Tsampas+19,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\bcp{Yes, going back to MIT days---we should
  include several more of these, if only to give readers the impression that
this is a well-studied mechanism (so formalizing its protections is
useful).}
\fi
%
The protections offered by such mechanisms are commonly described in terms
of concrete examples of attacks that they can prevent---corruption
of return addresses, buffer overflows, use of uninitialized variables,
etc.---leaving a more abstract characterization to the reader's intuition.
But these mechanisms can be quite intricate,
and they often require trading off protection and efficiency.
Thus, it would be useful to have a precise, generic, and formal
specification for stack
safety, both as a basis for comparing the security claims of different
enforcement techniques and for validating more rigorously that these claims
are met by particular implementations.

%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
The most abstract characterization stack safety that we are aware of is in
recent work by \citet{Skorstengaard+19}, which defines stack safety as the conjunction of two
properties:
{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
Informally, LSE says that the contents of
a caller's stack frame are not read and written while its callees (and their callees, etc.) are
executing, and WBCF says that callees always
return to the point just after their call (if they return at all).
Skorstengaard et al.{} formalize these properties by defining an idealized
machine in which they hold {\em by construction};
they then show that an enforcement
mechanism (in their case, hardware capabilities) guarantees stack safety, by
proving that an ordinary machine with enforcement can simulate the idealized machine.
%an enforcement mechanism (in their case, hardware
%capabilities)
%can then be shown to guarantee stack safety by exhibiting a fully abstract
%embedding function
%from the idealized machine to an ordinary machine equipped with this
%enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
%  understand that sentence.}
%%
%% While this work represents an important step in the right direction, we believe
%% there is still room to craft an operational characterization
%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
%% In other words, we seek formal criteria that can be applied
%% to a mechanism for protecting stack-based low-level programs to judge
%% whether or not it actually ``guarantees stack safety.''
%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
%% confused by the references to these types of properties, yet to be introduced?}
%% We begin with a strong
%% local state encapsulation property that closely models the behavior of eager
%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
%% is too strong to characterize stack safety; reasonable enforcement mechanisms
%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
%% to support this technique and argue that the new property, defined relative to an
%% arbitrary notion of observation. This property is still strong enough to prevent
%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
%% other reasonable enforcement mechanisms can enforce it.

In the present work, we also distinguish between protecting stack data (like
LSE) and
control flow (with WBCF).  But rather than defining stack safety as ``what
a certain idealized machine does,'' we formalize explicit criteria
based on the technical framework of language-based
security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
  survey?}\fi.
Our core properties are \emph{stepwise stack integrity} and
\emph{stepwise stack confidentiality}. WBCF is a desireable property, but
we can express safety in an information-flow sense without reference to it,
even safety from attacks that hijack control flow.  \bcp{On consideration
  (especially in light of the weird ``C confuses A by returning directly to
  A instead of returning to B so that B can finish up some stuff before
  returning to A'' example\footnote{Actually, this example is not so weird.
  If C, the attacker, can return a {\em result} directly to A, who is
  expecting a result from B, then that is clearly bad!}), I
agree now that we should discuss WBCF.  It should be kept as distinct as
possible from the ``data safety'' part, since it is not really an
information flow property.  But our treatment of it is interestingly
different from StkTokens, because of our call annotations, and we should
discuss that.}

@@

\section{Key Ideas by Example}
\label{sec:running-example}
\colorlet{lgray}{gray!40}
\colorlet{lred}{red!40}
\colorlet{lblue}{blue!20}

@@

\definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep}
if their adjacent pairs change together. We will represent this relation with
\(\machT \doteq \nachT\). \bcp{People may not be familiar with the notation}
\sna{It's a new notation, we use this enough that it seems worth picking an
  evocative symbol.}\bcp{Sorry -- I meant the notational convention of representing a
coinductive relation with inference rules in this way, not the symbol...}
We define it as a coinductive relation:

@@

\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \outside\).
\(\MPCT\) corresponds to the trace of a function from its call to its return.
Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
\(\nach \hookrightarrow \NPCT\langle f_\depth \rangle\). A system enjoys
{\em stepwise stack confidentiality} if for all such \(\MPCT\) and \(\NPCT\),
\(\MPCT \doteq \NPCT\).
